1.1 Task: Valid Variable Names.
Which of the following are valid/good variable names in R. What is wrong with the ones that are invalid/bad?

var1: valid

3way_handshake : Cannot begin with a number

.password : valid, will be a hidden variable

__test__ : cannot begin with an underscore

my-matrix-M : cannot contain hyphens

three.dimensional.array : valid

3D.distance : cannot begin with a number

.2objects : dot cannot be followed by a number

wz3gei92 : valid but inpractical

next : cannot use a reserved word

P : valid

Q : valid

R : valid

S : valid

T : can be used, but shouldn't because T also stands for TRUE

X : valid

is.larger? : cannot contain question mark


# 1.2 Task: Obscure Code.
**The code below works, but can be improved. Do improve it!**
```{r}
myIterAtoR.max <- 5
second_iterator.max<-7
col.NUM= 10
row.cnt =10


fwzy45 <- matrix(rep(1, col.NUM*row.cnt),nrow=row.cnt)

for(haystack in (2-1):col.NUM){
  for(needle in 1:row.cnt) {
    if(haystack>=myIterAtoR.max){
      fwzy45[haystack, needle]<-NA}
  }}



# Improved code
iter_max <- 5
col_num <- 10
row_num <- 10

A <- matrix(data = rep(x = 1, times = col_num*row_num), nrow = row_num)

for(i in 1:col_num){
  for(j in 1:row_num) {
    if (i >= iter_max) {
      A[i,j] <- NA
    } 
  }
}
```

# 1.3 Task: Better Formatting.
**Improve formatting and style of the following code:**

```{r}
simulate_genotype <- function( q, N=100 ) {
  if( length(q)==1 ){
    p <- (1 - q)
    f_gt <- c(p^2, 2*p*q, q^2) # AA, AB, BB
  }else{
    f_gt<-q
  }
  tmp <- sample( c('AA','AB','BB'), size =N, prob=f_gt, replace=T )
  return(tmp)
}

# Improved code
simulate_genotype <- function(q, N=100) {
  if (length(q) == 1) {
    p <- (1 - q)
    f_gt <- c(p^2, 2 * p * q, q^2) # AA, AB, BB
  }
  
  else {
    f_gt<-q
  }
  
  sample(c('AA','AB','BB'),
         size =N, 
         prob=f_gt, 
         replace=TRUE)
}
```

# 1.4 Task: Hidden Variable.
**Assign a vector of three last months (abbreviated in English) in a year to a hidden variable *my_months.**
```{r}

#rev(month.abb) gives dec-jan, [1:3] gives dec, nov, oct, rev(rev(month.abb)[1:3]) gives oct, nov, dec
.my_months <- rev(rev(month.abb)[1:3])

```

# 1.5 Task: Pipeline-friendly Function.
**Modify the function below so that it works with R pipes %>%:**

```{r}
library(tidyverse)

my_filter <- function(threshold = 1, x, scalar = 5) {
  x[x >= threshold] <- NA 
  x <- x * scalar
  return(x)
}

# Solution
my_filter <- function(x, threshold = 1, scalar = 5) {
  x[x >= threshold] <- NA 
  x * scalar
}

c(-5, 5) %>% my_filter()
```

# 1.6 Task: Untidy Code?
**Is the code below correct? Can it be improved?**

```{r}
simulate_phenotype <- function(pop_params, gp_map, gtype) {
  N = 3
  pop_mean <- pop_params[1]
  pop_var <- pop_params[2]
  pheno <- rnorm(n = N, mean = pop_mean, sd = sqrt(pop_var))
  effect <- rep(0, times = length(N))
  
  for (gt_iter in c('AA', 'AB', 'BB')) {
    effect[gtype == gt_iter] <- rnorm(n = sum(gtype == gt_iter), 
                                      mean = gp_map[gt_iter, 'mean_eff'], 
                                      sd = sqrt(gp_map[gt_iter, 'var_eff']))
  }
  dat <- data.frame(gt = gtype, raw_pheno = pheno, effect = effect, pheno = pheno + effect)
  return(dat)
}

simulate_phenotype(c(10,10),c(10,10),10)
```


# 2.1 Task: Computing Variance.
**Write a modular code (function or functions) that computes the sample standard deviation given a vector of numbers.** 
**Decide how to logically structure the code. Assume there are no built-in R functions for computing mean and variance** **available.

```{r}
sample_mean <- function(x) {
  sum(x) / length(x)
}

sum_squared_deviations <- function(x) {
  tmp <- (x - sample_mean(x))^2
  sum(tmp)
}

variance <- function(x) {
  sum_squared_deviations(x) / (length(x) -1)
}

std_dev <- function(x) {
  sqrt(variance(x))
}

sample_mean(c(1,2,3,4,5))

sum_squared_deviations(c(1,2,3,4,5))

variance(c(1,2,3,4,5))

std_dev(c(1,2,3,4,5))



```

 

